On Extending a Full-Sharing Design with Batched Scheduling

### 2015-04-15 paper.tex
performance on worst case scenarios

We now present experimental results about the usage of the batched
scheduling on the \emph{NS}, \emph{SS} and \emph{FS} designs. For the
sake of simplicity, for the \emph{SS} and \emph{FS} designs, we will be
presenting only the results for the lock free \emph{LF2} proposal
(\emph{SS}$_{LF_2}$ and \emph{FS}$_{LF_2}$), since they were the ones
that presented the lowest overheads in the previous chapters. For the
\emph{FS}$_{LF_2}$ design, we will use it with the \emph{PCC} procedure
enabled.

Concerning the benchmarks, we will be using the same five sets of
benchmarks presented before with the same number of runs per
benchmark, the same formula to calculate the overhead ratios, and the
same worst case scenario approach, where all threads begin with the
same query goal. To put the results in perspective, we experimented
with $1$, $8$, $16$, $24$ and $32$ threads (the maximum number of
cores available in our machine) with batched and local scheduling.

Table~\ref{tab_batched_overhead} shows the overhead ratios, when
compared with the \emph{NS} design with 1 thread (running with local
scheduling, PtMalloc and without TabMalloc), for the \emph{NS},
\emph{SS}$_{LF_2}$ and \emph{FS}$_{LF_2+PCC}$ designs (all running with
TabMalloc and TcMalloc), when running 1, 8, 16, 24 and 32 threads with
local and batched scheduling on the five sets of benchmarks (the
results by set of benchmark can be seen in the
Appendix.~\ref{app:batched}). For each design, the table has then two
columns, a column with \textbf{Local} that shows results already
presented in previous chapters for the local scheduling and a column
with \textbf{Batched} that shows the new results with the batched
scheduling. The overhead results presented in both \textbf{Local} and
\textbf{Batched} columns use as base time the execution times
presented in the \emph{NS} column of the Table~\ref{tab_benchs}.

\begin{table}[!ht]
\centering
\caption{Overhead ratios, when compared with the NS design with $1$
  thread (running with local scheduling, PtMalloc and without
  TabMalloc) for the NS, SS$_{LF_2}$, FS$_{LF_2+PCC}$ designs (with
  TabMalloc and TcMalloc), when running 1, 8, 16, 24 and 32 threads
  with local and batched scheduling on the five sets of benchmarks
  (best ratios by row and by design for the Minimum, Average and
  Maximum are in bold)}
%\scalebox{1.00} {
\begin{tabular}{ll|rr|rr|rr}
\hline\hline
\multicolumn{2}{c|}{\multirow{2}{*}{\bf Threads}} &
\multicolumn{2}{c|}{\multirow{1}{*}{\bf NS}} &
\multicolumn{2}{c}{\multirow{1}{*}{\bf SS$_{LF_2}$}} & 
\multicolumn{2}{|c}{\multirow{1}{*}{\bf FS$_{LF_2+PCC}$}}\\ \cline{3-8}
& 
& \multicolumn{1}{c}{\bf Local}
& \multicolumn{1}{c}{\bf Batched}
& \multicolumn{1}{|c}{\bf Local}
& \multicolumn{1}{c}{\bf Batched}
& \multicolumn{1}{|c}{\bf Local}
& \multicolumn{1}{c}{\bf Batched}\\
\hline
\multirow{4}{*}{\bf 1}
& {\bf Min }& {\bf 0.53}& 0.55& {\bf 0.54}& 0.55& 1.01& {\bf 0.95}\\
& {\bf Avg }& {\bf 0.78}& 0.82& {\bf 0.84}& 0.90& {\bf 1.30}& 1.46\\
& {\bf Max }& 1.06& {\bf 1.05}& {\bf 1.04}& {\bf 1.04}& {\bf 1.76}& 2.33\\
& {\bf StD }& 0.15& 0.14& 0.17& 0.16& 0.22& 0.44\\
\hline
\multirow{4}{*}{\bf 8}
& {\bf Min }& 0.66& {\bf 0.63}& 0.66& {\bf 0.63}& 1.16&{\bf  0.99}\\
& {\bf Avg }& {\bf 0.85}& 0.88& {\bf 0.92}& 0.93& {\bf 1.88}& 1.95\\
& {\bf Max }& {\bf 1.12}& 1.14& 1.20& {\bf 1.15}& {\bf 2.82}& 3.49\\
& {\bf StD }& 0.13& 0.14& 0.15& 0.14& 0.60& 0.79\\
\hline
\multirow{4}{*}{\bf 16}
& {\bf Min }& 0.85& {\bf 0.75}& 0.82& {\bf 0.77}& 1.17& {\bf 1.06}\\
& {\bf Avg }& {\bf 0.98}& 1.00& {\bf 1.04}& 1.05& {\bf 1.97}& 2.08\\
& {\bf Max }& {\bf 1.16}& 1.31& 1.31& {\bf 1.28}& {\bf 3.14}& 3.69\\
& {\bf StD }& 0.09& 0.17& 0.12& 0.13& 0.65& 0.83\\
\hline
\multirow{4}{*}{\bf 24}
& {\bf Min }& {\bf 0.91}& 0.93& 1.02& {\bf 0.98}& 1.16& {\bf 1.09}\\
& {\bf Avg }& {\bf 1.15}& 1.16& 1.22& {\bf 1.19}& {\bf 2.06}& 2.19\\
& {\bf Max }& 1.72& {\bf 1.60}& 1.81& {\bf 1.61}& {\bf 3.49}& 4.08\\
& {\bf StD }& 0.20& 0.21& 0.18& 0.16& 0.70& 0.91\\
\hline
\multirow{4}{*}{\bf 32}
& {\bf Min }& 1.05& {\bf 1.04}& {\bf 1.07}& 1.12& 1.33& {\bf 1.26}\\
& {\bf Avg }& 1.51& {\bf 1.49}& 1.54& {\bf 1.51}& {\bf 2.24}& 2.41\\
& {\bf Max }& {\bf 2.52}& 2.63& {\bf 2.52}& 2.62& {\bf 3.71}& 4.51\\
& {\bf StD }& 0.45& 0.45& 0.42& 0.43& 0.74& 1.02\\
\hline\hline
\end{tabular}%}
\label{tab_batched_overhead}
\end{table}

By observing Table~\ref{tab_batched_overhead}, we can see that, for
one thread, on average, local scheduling is sightly better than
batched on the three designs. For the \emph{NS} design we have $0.78$
and $0.82$, for the \emph{SS}$_{LF_2}$ design we have $0.84$ and $0.90$
and for the \emph{FS}$_{LF_2+PCC}$ design we have $1.30$ and $1.46$
average overhead ratios, for the local and batched scheduling
strategies, respectively.

As we scale the number of threads, one can observe that, for the
\emph{NS} and \emph{SS}$_{LF_2}$ designs both scheduling strategies have
similar minimum, average and maximum overhead ratios. For the
\emph{FS}$_{LF_2+PCC}$ design, the best minimum overhead ratio is
always for batched scheduling. The reader can observe on
Appendix~\ref{app:batched} that the minimum overhead values for $8$,
$16$, $24$ and $32$ threads are given by the benchmark belonging to
the model checking set (see Table~\ref{tab_benchs} for the
characteristics of the benchmarks). For the average and maximum
overhead ratio, local scheduling is always better than batched
scheduling. The reader can observe on Appendix.~\ref{app:batched} that
the maximum overhead values for $8$, $16$, $24$ and $32$ threads are
given by the pyramid benchmark in the path right set.

In summary, we can say that both the local and batched scheduling
strategies have similar overhead results on worst case scenarios for
the \emph{NS}, \emph{SS}$_{LF_2}$ and \emph{FS}$_{LF_2+PCC}$ designs.

@@

\section*{Acknowledgments}

>>>>

We now present experimental results about the usage of the batched
scheduling on the \emph{NS}, \emph{SS} and \emph{FS} designs. The
environment for our experiments was a machine with 32-Core AMD Opteron
(TM) Processor 6274 (2 sockets with 16 cores each) with 32G of main
memory, each processor with caches $L1$, $L2$ and $L3$ respectively
with the sizes of 64K, 2048K and 6144K, running the Linux kernel is
the 3.16.7-200.fc20.x86\_64, with Yap 6.3 compiled with gcc 4.8.

We used five sets of benchmarks. The \emph{Large Joins} and
\emph{WordNet} sets were obtained from the OpenRuleBench
project~\cite{Liang-09}; the \emph{Model Checking} set includes three
different specifications and transition relation graphs usually used
in model checking applications; the \emph{Path Left} and \emph{Path
  Right} sets implement two recursive definitions of the well-known
$path/2$ predicate, that computes the transitive closure in a graph,
using several different configurations of $edge/2$ facts. We
experimented graphs with the \emph{BTree} configuration with depth 17,
the \emph{Pyramid} and \emph{Cycle} configurations with depth 2000 and
the \emph{Grid} configuration with depth 35.

In order to have a deeper insight on the behavior of each benchmark,
and therefore clarify some of the results that are presented next, we
first characterize the benchmarks. The columns in
Table~\ref{tab_benchs} have the following meaning: (i) {\bf calls:} is
the number of different calls to tabled subgoals. It corresponds to
the number of paths in the subgoal tries. (ii) {\bf trie nodes:} is
the total number of trie nodes allocated in the corresponding
subgoal/answer trie structures. (iii) {\bf trie depth:} is the
minimum/average/maximum number of trie node levels required to
represent a path in the corresponding subgoal/answer trie
structures. Trie structures with smaller average values are more
amenable to higher lock contention. (iv) {\bf unique:} is the number
of different tabled answers found. It corresponds to the number of
paths in the answer tries. (v) {\bf repeated:} is the number of
redundant tabled answers found. (vi) {\bf NS:} is the average
execution time, in seconds, of ten runs for 1 thread with the NS
design without the TabMalloc memory allocator. In what follows, we
will use these times as the base times when computing the overhead
ratios for the other designs.

By observing Table~\ref{tab_benchs}, the \emph{Mondial} benchmark,
from the \emph{Large Joins} set, and the three \emph{Model
  Checking} benchmarks seem to be the benchmarks least amenable to
lock contention since they are the ones that find less unique answers
and that have the deepest trie structures. In this regard, the
\emph{Path Left} and \emph{Path Right} sets correspond to the
opposite case. They find a huge number of answers and have very
shallow trie structures. On the other hand, the \emph{WordNet} and
\emph{Path Right} sets have the benchmarks with the largest number
of different subgoal calls, which can reduce the probability of lock
contention because answers can be found for different subgoal calls
and therefore be inserted with minimum overlap. On the opposite side
are the \emph{Join2} benchmark, from the \emph{Large Joins} set,
and the \emph{Path Left} benchmarks, which have only a single tabled
subgoal call.

\begin{table}[ht]
\centering
\caption{Characteristics of the benchmark programs}
\scalebox{0.82} { 
\begin{tabular}{lrrccrrrcc}
\hline\hline
\multirow{2}{*}{\bf Bench}
& \multicolumn{3}{c}{\bf Tabled Subgoals}
&
& \multicolumn{4}{c}{\bf Tabled Answers}
& \multicolumn{1}{c}{\bf Time (sec)} \\ \cline{2-4}\cline{6-9}
& \multicolumn{1}{c}{\bf calls}
& \multicolumn{1}{c}{\bf trie nodes}
& \multicolumn{1}{c}{\bf trie depth}
&
& \multicolumn{1}{c}{\bf unique}
& \multicolumn{1}{c}{\bf repeated}
& \multicolumn{1}{c}{\bf trie nodes}
& \multicolumn{1}{c}{\bf trie depth}
& \multicolumn{1}{c}{\bf NS} \\
\hline
\multicolumn{10}{l}{\bf Large Joins} \\
\bf{Join2}    &       1 &       6 & 5/5/5 && 2,476,099 &         0 &
2,613,660 &    5/5/5 &  2.85 \\
\bf{Mondial}  &      35 &      42 & 3/4/4 &&     2,664 & 2,452,890 &
   14,334 &    6/7/7 &  0.84 \\
\hline
\multicolumn{10}{l}{\bf WordNet} \\
\bf{Clusters} & 117,659 & 235,319 & 2/2/2 &&   166,877 &   161,853 &
284,536 &    1/1/1 &   0.83 \\
\bf{Holo}     & 117,657 & 235,315 & 2/2/2 &&    74,838 &        54 &
192,495 &    1/1/1 &   0.75 \\
\bf{Hyper}    & 117,657 & 235,315 & 2/2/2 &&   698,472 &     8,658 &
816,129 &    1/1/1 &   1.42 \\
\bf{Hypo}     & 117,657 & 117,659 & 2/2/2 &&   698,472 &    20,341 &
816,129 &    1/1/1 &   1.53 \\
\bf{Mero}     & 117,657 & 117,659 & 2/2/2 &&    74,838 &        13 &
192,495 &    1/1/1 &   0.74 \\
\bf{Tropo}    & 117,657 & 235,315 & 2/2/2 &&       472 &         0 &
118,129 &    1/1/1 &   0.66 \\
\hline
\multicolumn{10}{l}{\bf Model Checking} \\
\bf{IProto}   &       1 &       6 & 5/5/5 &&   134,361 &   385,423 &
1,554,896 &  4/51/67 & 2.70 \\
\bf{Leader}   &       1 &       5 & 4/4/4 &&     1,728 &   574,786 &
   41,788 & 15/80/97 & 3.51  \\
\bf{Sieve}    &       1 &       7 & 6/6/6 &&       380 & 1,386,181 &
    8,624 & 21/53/58 & 18.50 \\
\hline
\multicolumn{10}{l}{\bf Path Left} \\
\bf{BTree}    &       1 &       3 & 2/2/2 && 1,966,082 &         0 &
2,031,618 &    2/2/2 &   1.53 \\
\bf{Cycle}    &       1 &       3 & 2/2/2 && 4,000,000 &     2,000 &
4,002,001 &    2/2/2 &   3.52 \\
\bf{Grid}     &       1 &       3 & 2/2/2 && 1,500,625 & 4,335,135 &
1,501,851 &    2/2/2 &  1.93 \\
\bf{Pyramid}  &       1 &       3 & 2/2/2 && 3,374,250 & 1,124,250 &
3,377,250 &    2/2/2 &   3.08 \\
\hline
\multicolumn{10}{l}{\bf Path Right} \\
\bf{BTree}    & 131,071 & 262,143 & 2/2/2 && 3,801,094 &         0 &
3,997,700 &    1/2/2 &  2.33 \\
\bf{Cycle}    &   2,001 &   4,003 & 2/2/2 && 8,000,000 &     4,000 &
8,004,001 &    1/2/2 &  3.55 \\
\bf{Grid}     &   1,226 &   2,453 & 2/2/2 && 3,001,250 & 8,670,270 &
3,003,701 &    1/2/2 &  2.32 \\
\bf{Pyramid}  &   3,000 &   6,001 & 2/2/2 && 6,745,501 & 2,247,001 &
6,751,500 &    1/2/2 &  3.17 \\
\hline\hline
\end{tabular}}
\label{tab_benchs}
\end{table}

Note that for the moment our goal is to evaluate the robustness of our
implementation when exposed to \emph{worst case} scenarios. By
focusing on \emph{worst case} scenarios, we show the lowest bounds in
terms of performance that each design might achieve when applied/used
with other real world applications/programs. Moreover, by testing the
framework with worst case scenarios, we avoid the peculiarities of the
program at hand and we try to focus on measuring the real value of our
designs. Thus, we will follow a common approach to create worst case
scenarios and we will run all threads starting with the same query
goal. By doing this, it is expected that all threads will access the
table space, to check/insert for subgoals and answers, at similar
times, thus causing a huge stress on the same critical regions. To put
the results in perspective, we experimented with $1$, $8$, $16$, $24$
and $32$ threads (the maximum number of cores available in our
machine) with local and batched scheduling on the five sets of
benchmarks presented in the early subsection.

Table~\ref{tab_batched_overhead} shows the overhead ratios, when
compared with the NS design with 1 thread (running with local
scheduling), for the NS, SS and FS designs (all running TabMalloc),
when running 1, 8, 16, 24 and 32 threads with local and batched
scheduling on the five sets of benchmarks. For each design, the table
has then two columns, a column with \emph{Local} that shows results
for the local scheduling and a column with \emph{Batched} with results
for batched scheduling. The overhead results presented in both
\emph{Local} and \emph{Batched} columns use as base time the execution
times presented in the \emph{NS} column of the Table~\ref{tab_benchs}.

\begin{table}[!ht]
\centering
\caption{Overhead ratios, when compared with the NS design with $1$
  thread (running with local scheduling without TabMalloc) for the NS,
  SS, FS designs (with TabMalloc), when running 1, 8, 16, 24 and 32
  threads with local and batched scheduling on the five sets of
  benchmarks (best ratios by row and by design for the Minimum,
  Average and Maximum are in bold)}
%\scalebox{1.00} {
\begin{tabular}{ll|cc|cc|cc}
\hline\hline
\multicolumn{2}{c|}{\multirow{2}{*}{\bf Threads}} &
\multicolumn{2}{c|}{\multirow{1}{*}{\bf NS}} &
\multicolumn{2}{c}{\multirow{1}{*}{\bf SS}} & 
\multicolumn{2}{|c}{\multirow{1}{*}{\bf FS}}\\ \cline{3-8}
& 
& \multicolumn{1}{c}{\bf Local}
& \multicolumn{1}{c}{\bf Batched}
& \multicolumn{1}{|c}{\bf Local}
& \multicolumn{1}{c}{\bf Batched}
& \multicolumn{1}{|c}{\bf Local}
& \multicolumn{1}{c}{\bf Batched}\\
\hline
\multirow{4}{*}{\bf 1}
& {\bf Min }& {\bf 0.53}& 0.55& {\bf 0.54}& 0.55& 1.01& {\bf 0.95}\\
& {\bf Avg }& {\bf 0.78}& 0.82& {\bf 0.84}& 0.90& {\bf 1.30}& 1.46\\
& {\bf Max }& 1.06& {\bf 1.05}& {\bf 1.04}& {\bf 1.04}& {\bf 1.76}& 2.33\\
& {\bf StD }& 0.15& 0.14& 0.17& 0.16& 0.22& 0.44\\
\hline
\multirow{4}{*}{\bf 8}
& {\bf Min }& 0.66& {\bf 0.63}& 0.66& {\bf 0.63}& 1.16&{\bf  0.99}\\
& {\bf Avg }& {\bf 0.85}& 0.88& {\bf 0.92}& 0.93& {\bf 1.88}& 1.95\\
& {\bf Max }& {\bf 1.12}& 1.14& 1.20& {\bf 1.15}& {\bf 2.82}& 3.49\\
& {\bf StD }& 0.13& 0.14& 0.15& 0.14& 0.60& 0.79\\
\hline
\multirow{4}{*}{\bf 16}
& {\bf Min }& 0.85& {\bf 0.75}& 0.82& {\bf 0.77}& 1.17& {\bf 1.06}\\
& {\bf Avg }& {\bf 0.98}& 1.00& {\bf 1.04}& 1.05& {\bf 1.97}& 2.08\\
& {\bf Max }& {\bf 1.16}& 1.31& 1.31& {\bf 1.28}& {\bf 3.14}& 3.69\\
& {\bf StD }& 0.09& 0.17& 0.12& 0.13& 0.65& 0.83\\
\hline
\multirow{4}{*}{\bf 24}
& {\bf Min }& {\bf 0.91}& 0.93& 1.02& {\bf 0.98}& 1.16& {\bf 1.09}\\
& {\bf Avg }& {\bf 1.15}& 1.16& 1.22& {\bf 1.19}& {\bf 2.06}& 2.19\\
& {\bf Max }& 1.72& {\bf 1.60}& 1.81& {\bf 1.61}& {\bf 3.49}& 4.08\\
& {\bf StD }& 0.20& 0.21& 0.18& 0.16& 0.70& 0.91\\
\hline
\multirow{4}{*}{\bf 32}
& {\bf Min }& 1.05& {\bf 1.04}& {\bf 1.07}& 1.12& 1.33& {\bf 1.26}\\
& {\bf Avg }& 1.51& {\bf 1.49}& 1.54& {\bf 1.51}& {\bf 2.24}& 2.41\\
& {\bf Max }& {\bf 2.52}& 2.63& {\bf 2.52}& 2.62& {\bf 3.71}& 4.51\\
& {\bf StD }& 0.45& 0.45& 0.42& 0.43& 0.74& 1.02\\
\hline\hline
\end{tabular}%}
\label{tab_batched_overhead}
\end{table}

By observing Table~\ref{tab_batched_overhead}, we can see that, for
one thread, on average, local scheduling is sightly better than
batched on the three designs. For the NS design we have $0.78$ and
$0.82$, for the SS design we have $0.84$ and $0.90$ and for the FS
design we have $1.30$ and $1.46$ average overhead ratios, for the
local and batched scheduling strategies, respectively. As we scale the
number of threads, one can observe that, for the NS and SS designs
both scheduling strategies have similar minimum, average and maximum
overhead ratios. For the FS design, the best minimum overhead ratio is
always for batched scheduling. For the average and maximum overhead
ratio, local scheduling is always better than batched scheduling. In
summary, we can say that both the local and batched scheduling
strategies have similar overhead results on worst case scenarios for
the NS, SS and FS designs.

@@

%\section*{Acknowledgments}
